---
title: "Z Score QA"
output: html_notebook
---

Notes:
Instruction of Calculator :	
	1.  copy raw ratios to column B (Raw Ratio)
	2.  Final wonsorized raw ratio and its z-score is returned in column L amd M. Use only these two columns as the final result, not the iteration 4
	3.  Cell O3 shows how many iterations the final result runs
	
Algorithm:	
	1. Set Iteration =1;
	2. Compute z-scores (cross-sectional) for all the raw ratios in the base universe selected
	1.1 If All |z-score|<5, use this z-score as the final result. Stop iterative process.
	1.2 If Any |z-score| >=5, then flag any instance that has a z-score > |3| as outlier 
	    
	3. Compute a new_mean and new_stdev of raw ratios from #1 that are not flagged as outliers (having a z-score with z-scores <= |3|)
	
	4. For any raw ratios that,
	
	
	
	5. Iteration = Iteration + 1;
	6. Repeat steps 2 to 5. End process at 2.1 or until Iteration = 4, whichever is earliest.
	
	(This iteration number is slightly off from the SQL s.p. iteration column. )

```{r}
library(dplyr)
```

1. build a z score function
```{r}
zscore = function(data) {
  z = (data-mean(data))/sd(data)
  return(z)
}

# test
# a = c(1:10)
# zscore(a)
```

2. build a winsorize function
```{r}
# winsorize test code

a = as.double(c(1:49))
a = c(a,1E+16)
sigma = 1 # times of sigma check

a[abs(zscore(a)) <= 3]

mean(a[abs(zscore(a)) <= 3])
sd(a[abs(zscore(a)) <= 3])


b = case_when(
  zscore(a) <= -3 ~ mean(a[abs(zscore(a)) <= 3]) - 3*sd(a[abs(zscore(a)) <= 3]),
  zscore(a) >= 3 ~ mean(a[abs(zscore(a)) <= 3]) + 3*sd(a[abs(zscore(a)) <= 3]),
  TRUE ~ a
)

summary(b)
zscore(b)

c = data.frame(a,b)

c[c$b == max(c$b),]
td$test2[297]

# function it up
winsorize = function(data = NULL, sigAdj=NULL) {
  require(dplyr)
  
  new_data = data[abs(zscore(data)) <= sigAdj]
  new_mean = mean(new_data)
  new_sd = sd(new_data)
  lowerb = new_mean - sigAdj*new_sd
  upperb = new_mean + sigAdj*new_sd
  
  wd = case_when(
  data < lowerb ~ lowerb,
  data > upperb ~ upperb,
  TRUE ~ data
  )
  
  return(wd)
}

winsorize1 = function(data = NULL, sigAdj=NULL) {
  require(dplyr)
  
  new_data = data[abs(zscore(data)) <= sigAdj]
  new_mean = mean(new_data)
  new_sd = sd(new_data)
  lowerb = new_mean - sigAdj*new_sd
  upperb = new_mean + sigAdj*new_sd
  
  wd = case_when(
  zscore(data) <= -3 ~ lowerb,
  zscore(data) >= -3 ~ upperb,
  TRUE ~ data
  )
  
  return(wd)
}

# test function
a = as.double(c(1:49))
a = c(a,1E+16)
results = data.frame(
  "raw value" = a, 
  "zscore" = zscore(a),
  "wsValue" = winsorize(a, 3),
  "wsZ" = zscore(winsorize(a, 3)),
  "wsValueAlt" = winsorize1(a, 3),
  "wsZAlt" = zscore(winsorize1(a, 3))
  )
results



```

3. build a while loop function
```{r}
# a = as.double(c(1:100))
a = as.double(c(1:9))
a = c(a,1E+16)
a = td$test2

sc = 5 # times of sigma for checking outlier
sa = 3 # times of sigma for adjusting raw value
maxitr = 100 # max iteration limit (use equal sigCheck & sigAdj to test)

i = 1 # initialization parameter
winsCt = 0 # counts of winsorization
maz = max(abs(zscore(a)))

while (i == 1) {
  if ( maz > sc & winsCt <= maxitr) {
    a = winsorize(a,sa)
    maz = max(abs(zscore(a)))
    winsCt = winsCt+1
  } else {
    i = 0
    # a = a
    # maz = max(abs(zscore(a)))
  }
  
  print(max(a))
  print(maz)
  print(i)
  # using 
}

rm(list = c("sigAdj","sigCheck","sigma","winsCt","maz","maxitr","i"))

####### function it up
wins_rs = function(data=NULL, sigCheck=NULL, sigAdj=NULL) {
  # define additional parameters
  maxitr = 100 # max iteration limit (use equal sigCheck & sigAdj to test)
  i = 1 # initialization parameter
  wCt = 0 # counts of winsorization
  zsLog = zscore(data)
  dLog = data
  
  # while loop
  while (i == 1) {
    maz = max(abs(zscore(data)))
    if ( maz > sigCheck && wCt <= maxitr) {
      i = 1
      data = winsorize(data,sigAdj)
      # maz = max(abs(zscore(data)))
      zsLog = data.frame(zsLog,zscore(data))
      dLog = data.frame(dLog, data)
      wCt = wCt+1
    } else {
      i = 0
      # data = data
      # maz = max(abs(zscore(data)))
      # zsMx = data.frame(zsMx,zscore(data))
      # wCt = wCt
    }
    
  }
    
  list.out = list("data" = data, 
                  "dataz" = zscore(data),
                  "WinsCount" = wCt,
                  "LastZscore" = maz,
                  "zlog" = zsLog,
                  "dlog" = dLog
                  )
  return(list.out)
}

# test
a = as.double(c(1:49))
a = c(a,1E+16)
op = wins_rs(a,5,3)

# op$data
# op$dataz
op$WinsCount
op$zlog
op$dlog

```

4. test using other values
```{r}

path = "C:/Users/patrick.fang/OneDrive - IHS Markit/ResearchSignals/Projects/20191226_ZScoreQA"
file = "testCases.txt"

td = read.table(file.path(path,file), header = TRUE, sep = "\t")

# rm(list = c("ouput1","ouput2","ouput3"))

output1 = wins_rs(td$test1,5,3)
output2 = wins_rs(td$test2,5,3)
output3 = wins_rs(td$test3,5,3)

output = data.frame("output1" = c(zscore(output1$data), output1$WinsCount),
                    "output2" = c(zscore(output2$data), output2$WinsCount),
                    "output3" = c(zscore(output3$data), output3$WinsCount)
)

write.table(output, file.path(path, "testOutputs.txt"), sep = '\t', row.names=FALSE, col.names = TRUE, quote = FALSE)



```

